---
title: "Data Reuse in Metabolomics"
author: "Analysis by Rachel Spicer, github:RASpicer"
date: "03/06/2018"
output: 
  html_document:
    code_folding: hide
    number_sections: yes
    theme: cerulean
    keep_md: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyr)
library(plotly)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy=TRUE, message = FALSE,fig.path = "figs/",fig.keep=TRUE)
```
# Data Processing 
This RMarkdown contains the code used for analysis for the subsection <b>Data Reuse in Metabolomics</b> of Chapter 3 of the thesis <b>Fit for purpose? A metascientific analysis of metabolomics data in public repositories</b>. 

## Functions
This sections contains functions that are used for processing the reuse data.
```{r functions}
# Functions
# Extract which studies have been reused
ReusedStudies <- function(x){
  Studies <- as.character(x$`Study(-ies)`)
  Studies <- unlist(strsplit(Studies, ","))
  # remove instances where a single study is not referenced
  Studies <- Studies[!grepl("-", Studies)]
  Studies <- Studies[!grepl("All", Studies)]
  Studies <- Studies[!grepl("GNPS", Studies)]
  return(Studies)
}

# Produce an ordered frequency table of the number of reuses per study
UniqueStudiesCount <- function(x){
  Studies <- ReusedStudies(x)
  # Produce a frequency table of the number of reuses per study
  UniqueStudiesCount <- as.data.frame(table(Studies))
  # Order by Frequency of reuse
  UniqueStudiesCountOrdered <- UniqueStudiesCount[order(-UniqueStudiesCount$Freq),]
  return(UniqueStudiesCountOrdered)
}

#  Find the count of MetaboLights studies being reused
MetaboLightsReuse <- function(x){
  MetaboLightsC <- x[grepl("MTBLS", x)]
  MetaboLightsCount <- as.data.frame(table(MetaboLightsC))
  MetaboLightsCount <- MetaboLightsCount[order(-MetaboLightsCount$Freq),]
  MetaboLightsCount$Repository <- rep("MetaboLights", nrow(MetaboLightsCount))
  colnames(MetaboLightsCount) <- c("Study", "Frequency", "Repository")
  return(MetaboLightsCount)
}

#  Find the count of Metabolomics Workbench studies being reused
MWReuse <- function(x){
  MetabolomicsWorkbenchC <- x[grepl("ST000", x)]
  MetabolomicsWorkbenchCount <- as.data.frame(table(MetabolomicsWorkbenchC))
  MetabolomicsWorkbenchCount <- MetabolomicsWorkbenchCount[order(-MetabolomicsWorkbenchCount$Freq),]
  MetabolomicsWorkbenchCount$Repository <- rep("Metabolomics Workbench", nrow(MetabolomicsWorkbenchCount))
  colnames(MetabolomicsWorkbenchCount) <- c("Study", "Frequency", "Repository")
  return(MetabolomicsWorkbenchCount)
}

#  Find the count of GNPS studies being reused
GNPSReuse <- function(x){
  GNPSC <-  x[grepl("MSV", x)]
  GNPSCount <- as.data.frame(table(GNPSC))
  GNPSCount <- GNPSCount[order(-GNPSCount$Freq),]
  GNPSCount$Repository <- rep("GNPS", nrow(GNPSCount))
  colnames(GNPSCount) <- c("Study", "Frequency", "Repository")
  return(GNPSCount)
}

# Combine the amount of reuse per repository into a single data frame
# Requires input of the number of GNPS, MetaboLights and Metabolomics Workbench studies which have not been reused
CombStudies <- function(Studylist, GNPS, ML, MW){
  GNPSs <- GNPSReuse(Studylist)
  MLs <- MetaboLightsReuse(Studylist)
  MWs <- MWReuse(Studylist)
  AllStudies <- rbind(GNPSs, MLs, MWs)
  MLnr <- data.frame(Study = rep("Study", ML), Frequency = rep(0, ML), Repository = rep('MetaboLights'))
  # MW 609 -  nrow(MetabolomicsWorkbenchCount) = 364
  MWnr <- data.frame(Study = rep("Study", MW), Frequency = rep(0, MW), Repository = rep('Metabolomics Workbench'))
  # GNPS 688 - nrow(GNPSCount) = 604
  GNPSnr <- data.frame(Study = rep("Study", GNPS), Frequency = rep(0, GNPS), Repository = rep('GNPS'))
  # Combine into single dataframe
  AllStudies <- rbind(AllStudies, MLnr, MWnr, GNPSnr)
  AllStudies$Repository <- as.factor(AllStudies$Repository)
  levels(AllStudies$Repository) <- gsub(" ", "\n", levels(AllStudies$Repository))
  return(AllStudies)
}

# Create data frame of the percentage of reuse at each frequency
PerReuse <- function(x){
  # The number of studies with each frequency of reuse
  Per0 <- nrow(x[x$Frequency == 0,])/nrow(x) * 100
  Per1 <- nrow(x[x$Frequency == 1,])/nrow(x) * 100
  Per2 <- nrow(x[x$Frequency == 2,])/nrow(x) * 100
  Per3 <- nrow(x[x$Frequency == 3,])/nrow(x) * 100
  Per4 <- nrow(x[x$Frequency == 4,])/nrow(x) * 100
  df <- data.frame("FrequencyofReuse" = c(0,1,2,3,4), "Percentage" = c(Per0,Per1,Per2,Per3,Per4))
  return(df)
}

# Create a data frame of the reuse per repository and the reuse across repositories
PerReuseRepository <- function(x){
  GNPS <- x[x$Repository == "GNPS",]
  ML <- x[x$Repository == "MetaboLights",]
  MW <- x[x$Repository == "Metabolomics\nWorkbench",]
  GNPSPer <- PerReuse(GNPS)
  MLPer <- PerReuse(ML)
  MWPer <- PerReuse(MW)
  AllPer <- PerReuse(x)
  df = cbind(GNPSPer, MLPer[,2], MWPer[,2], AllPer[,2])
  colnames(df) <- c("Frequency", "GNPS", "MetaboLights", "Metabolomics Workbench","All Studies")
  return(df)
}

# Percentage of studies published in each year that have been reused
# Requires input of list of studies that have been reused and all studies in repository
AgeReuse <- function(Repos, ReStud){
  # Reused studies
  Reused <- Repos[Repos$StudyID %in% ReStud$Studies ,]
  # Not reused studies
  Not <- Repos[!Repos$StudyID %in% ReStud$Studies ,]
  # Publication Year of studies that have been reused
  ReusedAge <- as.data.frame(table(Reused$Year))
  # Publication Year of studies that have not been reused
  NotAge <- as.data.frame(table(Not$Year))
  df <- cbind(ReusedAge,NotAge[1:nrow(ReusedAge),2])
  colnames(df) <- c("Year", "Reused", "Not")
  # Calculate percentage of studies reused per year
  df$Percentage <- round(df$Reused/(df$Reused+df$Not)*100, 2)
  # Remove frequency columns
  df <- df[,-(2:3)]
  return(df)
}
```

## Reuse per Year

Calculate the number of publications that reuse data from each repository per year.
```{r dataprocessing}
# Load Reuse data
DataReuse <- read.csv("../data/DataReuseMetabolomics.csv", check.names = FALSE)
DataReuse$Year <- as.factor(DataReuse$Year)

# Convert data re-use per year to a factor (including 0 for the year 2012)
DataReuse$Year <- factor(DataReuse$Year, levels = c("2012",levels(DataReuse$Year)))
# Find the frequency of data reuse per year
DataReuseYear <- as.data.frame(table(DataReuse$Year))
colnames(DataReuseYear) <- c("Year","Frequency")
DataReuseYear$Repository <- rep("All")

# Count number of reuses per repository
# MetaboLights
MLNo <- DataReuse[grepl("MetaboLights", DataReuse$Repository),]
# Metabolomics Workbench
MWNo <- DataReuse[grepl("Metabolomics Workbench", DataReuse$Repository),]
# GNPS
GNPSNo <-  DataReuse[grepl("GNPS", DataReuse$Repository),]

# Create contigency tables of the frequency of reuse per year
# adding a level for the year prior to any examples of reuse for plotting
# MetaboLights
MLYear <- as.data.frame(table(MLNo$Year))
colnames(MLYear) <- c("Year","Frequency")
MLYear$Repository <- rep("MetaboLights")
MLYear$Year <- as.factor(MLYear$Year)
MLYear$Year <- factor(MLYear$Year, levels = c("2012",levels(MLYear$Year)))
# Metabolomics Workbench
MWYear <- as.data.frame(table(MWNo$Year))
colnames(MWYear) <- c("Year","Frequency")
MWYear$Repository <- rep("Metabolomics Workbench")
MWYear$Year <- as.factor(MWYear$Year)
MWYear$Year <- factor(MWYear$Year, levels = c("2016",levels(MWYear$Year)))
# GNPS
GNPSYear <- as.data.frame(table(GNPSNo$Year))
colnames(GNPSYear) <- c("Year","Frequency")
GNPSYear$Repository <- rep("GNPS")
GNPSYear$Year <- as.factor(GNPSYear$Year)
GNPSYear$Year <- factor(GNPSYear$Year, levels = c("2015",levels(GNPSYear$Year)))

# Recombine data into single data frame
Freqperyear <- rbind(DataReuseYear, MLYear, MWYear, GNPSYear)
Freqperyear$Repository <- as.factor(Freqperyear$Repository)

# Remove 2018 for ploting
Y2018 <- grepl("2018",Freqperyear$Year)
No2018 <- Freqperyear[!Y2018,]
```

## Reuse per Study

Calculate the frequency of reuse of each study.

```{r reuseperstudy}
# Find the total number of studies being reused (including reuse in Spicer et al. 2018)
TotalUniqueStudies <- ReusedStudies(DataReuse)

# As of 15/2/18 there are 329 MetaboLghts studies, 614 Metabolomics Workbench studies and 691 GNPS studies
# GNPS 691 - nrow(GNPSCount) = 606
# ML 332 - nrow(MetaboLightsCount) = 187
# MW 614 -  nrow(MWCount) = 369
AllStudies <- CombStudies(TotalUniqueStudies, 606, 187, 369)

# Find the percentage reuse at each frequency across all repositories
Reuseper <- PerReuseRepository(AllStudies)
# melt data for ploting
Reuseperg <- melt(Reuseper, "Frequency")
colnames(Reuseperg) <- c("Frequency", "Repository","Percentage")
Reuseperg$Frequency <- factor(Reuseperg$Frequency, levels = c("4","3","2","1","0"))
levels(Reuseperg$Repository) <- gsub(" ", "\n", levels(Reuseperg$Repository))

# Find the total number of studies being reused (excluding reuse in Spicer et al. 2018)
TotalUniqueStudiesnoSpicer <- ReusedStudies(DataReuse[!grepl("Spicer", DataReuse$`Author(-s)`),])

# As of 15/2/18 there are 329 MetaboLghts studies, 614 Metabolomics Workbench studies and 691 GNPS studies
# GNPS 691 - nrow(GNPSCountnoSpicer) = 606
# ML 329 - nrow(MetaboLightsCountnoSpicer) = 275
# MW 614 -  nrow(MWCountnoSpicer) = 592
AllStudiesnospicer <- CombStudies(TotalUniqueStudiesnoSpicer, 606, 275, 592)

# Find the percentage reuse at each frequency across all repositories
Reusepernospicer <- PerReuseRepository(AllStudiesnospicer)
# melt data for ploting
Reusepernospicerg <- melt(Reusepernospicer, "Frequency")
colnames(Reusepernospicerg) <- c("Frequency", "Repository","Percentage")
Reusepernospicerg$Frequency <- factor(Reusepernospicerg$Frequency, levels = c("4","3","2","1","0"))
levels(Reusepernospicerg$Repository) <- gsub(" ", "\n", levels(Reusepernospicerg$Repository))
```

## Same Authors

Studies were examined to see whether they reused data produced by the same set of authors, submitters or study owners. This is important as one of the most common reasons researchers cite for not sharing data is the fear that they will be able to generate less publications from their data, and other researchers will scoop them.

Percentage of studies that reuse data that share at least one author/submitter/study owner with the original study:
```{r sameauth}
# Calculate percentage of studies where the original authors have reused their own data
SameAuth <- round(sum(DataReuse$`Same Group`)/(nrow(DataReuse)) * 100, digits = 2)
SameAuth
```

## Time until Reuse
Calculate the time between data being made public and its reuse. The average number of years is:
```{r time}
# Load repository data 
# MetaboLights
ML <- read.csv("../data/MLStudiesTime.csv", check.names = FALSE, stringsAsFactors = FALSE)
ML$Year <- format(as.Date(ML$StudyPublicationDate,"%d/%m/%y"),"%Y")
# Metabolomics Workbench
MW <- read.csv("../data/MWStudiesTime.csv", check.names = FALSE, stringsAsFactors = FALSE)
MW$Year <- format(as.Date(MW$ReleaseDate,"%d/%m/%Y"),"%Y")
# GNPS
GNPS <- read.csv("../data/GNPSStudiesTime.csv", check.names = FALSE, stringsAsFactors = FALSE)
GNPS$Year <- format(as.Date(GNPS$`Upload Date`, "%b. %d,%Y"),"%Y")

# Split DataReuse by study, but maintain details of reuse
UniqueStudiesReuse <- DataReuse %>% 
  mutate(`Study(-ies)` = strsplit(as.character(`Study(-ies)`), ",")) %>% 
  unnest(`Study(-ies)`)
# rename colname `Study(-ies)` to StudyID for joining
colnames(UniqueStudiesReuse)[9] <- "StudyID"
# rename year to ReuseYear
colnames(UniqueStudiesReuse)[3] <- "ReuseYear"

# Create data frames of just StudyID and Year
MLSY <- as.data.frame(cbind(ML[,1], ML[,7]))
MWSY <- as.data.frame(cbind(MW[,1], MW[,10]))
GNPSSY <- as.data.frame(cbind(GNPS[,2], GNPS[,10]))
# join data frames
StIDY <- rbind(MLSY,MWSY,GNPSSY)
colnames(StIDY) <- c("StudyID", "Year")

# add year to studies
UniqueStudiesReuseYear <- left_join(UniqueStudiesReuse, StIDY, by = "StudyID")
# Convert years from factor to numeric
UniqueStudiesReuseYear$Year <- as.numeric(as.character(UniqueStudiesReuseYear$Year))
UniqueStudiesReuseYear$ReuseYear <- as.numeric(as.character(UniqueStudiesReuseYear$ReuseYear))
# Calculate time between data being published and reuse 
UniqueStudiesReuseYear$Time2Reuse <- UniqueStudiesReuseYear$ReuseYear - UniqueStudiesReuseYear$Year

# Create contigency table
Time2Reuse <- as.data.frame(table(UniqueStudiesReuseYear$Time2Reuse))
colnames(Time2Reuse) <- c("Years", "Frequency")
Time2Reuse$Percentage <- Time2Reuse$Frequency/sum(Time2Reuse$Frequency)*100

# Average time until reuse 
AveTime <- round(mean(UniqueStudiesReuseYear$Time2Reuse, na.rm = T), 2)
AveTime
```


## Time Data made Public

Calculate the time until data reuse per study in each repository.

```{r age}
# Unique Studies that have been reused + Frequency
AllStudiesReused <- UniqueStudiesCount(DataReuse)

# Percentage reused Metabolights studies
MLAge <- AgeReuse(ML,AllStudiesReused)

# Percentage reused Metabolomics Workbench studies
MWAge <- AgeReuse(MW,AllStudiesReused)

# Percentage reused Metabolomics Workbench studies
GNPSAge <- AgeReuse(GNPS,AllStudiesReused)

# Create contigency tables of the frequency of reuse per year
# adding a level for the year prior to any examples of reuse for plotting
# MetaboLights
MLAge$Repository <- rep("MetaboLights")
MLAge$Year <- as.factor(MLAge$Year)
# Metabolomics Workbench 
MWAge$Repository <- rep("Metabolomics Workbench")
MWAge$Year <- as.factor(MWAge$Year)
# GNPS
GNPSAge$Repository <- rep("GNPS")
GNPSAge$Year <- as.factor(GNPSAge$Year)

# Recombine data into single data frame
PerAge <- rbind(MLAge, MWAge, GNPSAge)
PerAge$Repository <- as.factor(PerAge$Repository)
# Add row showing 0 reuse for GNPS in 2017
GNPS2017 <- c("2017", 0.00, "GNPS")
PerAge <- rbind(PerAge,GNPS2017)
PerAge$Percentage <- as.numeric(PerAge$Percentage)

# Reuse excluding Spicer et al, (2017)
NoSpicerReused <- UniqueStudiesCount(DataReuse[!grepl("Spicer", DataReuse$`Author(-s)`),])

# Percentage reused Metabolights studies
MLNSAge <- AgeReuse(ML,NoSpicerReused)

# Percentage reused Metabolomics Workbench studies
MWNSAge <- AgeReuse(MW,NoSpicerReused)

# Percentage reused Metabolomics Workbench studies
GNPSNSAge <- AgeReuse(GNPS,NoSpicerReused)

# Create contigency tables of the frequency of reuse per year
# adding a level for the year prior to any examples of reuse for plotting
# MetaboLights
MLNSAge$Repository <- rep("MetaboLights")
MLNSAge$Year <- as.factor(MLNSAge$Year)
# Metabolomics Workbench 
MWNSAge$Repository <- rep("Metabolomics Workbench")
MWNSAge$Year <- as.factor(MWNSAge$Year)
# GNPS
GNPSNSAge$Repository <- rep("GNPS")
GNPSNSAge$Year <- as.factor(GNPSNSAge$Year)

# Recombine data into single data frame
PerNSAge <- rbind(MLNSAge, MWNSAge, GNPSNSAge)
PerNSAge$Repository <- as.factor(PerNSAge$Repository)
# Add row showing 0 reuse for Metabolomics Workbench and GNPS in 2017
MW2017 <- c("2017", 0.00, "Metabolomics Workbench")
PerNSAge <- rbind(PerNSAge,MW2017,GNPS2017)
PerNSAge$Percentage <- as.numeric(PerNSAge$Percentage)
```

Calculate the number of studies released per repository per year.

```{r perreuseperyear}
# Average deposition date of MetaboLights studies
MLSDT <- round(mean(as.numeric(ML$Year), na.rm = T), 2)

# Average deposition date of Metabolomics Workbench studies
MWSDT <- round(mean(as.numeric(MW$Year), na.rm = T), 2)

# Average deposition date of GNPS studies
GNPSDT <- round(mean(as.numeric(GNPS$Year), na.rm = T), 2)

# Average deposition date of all studies
Years <- cbind(ML$Year, MW$Year, GNPS$Year)
YearsDT <- round(mean(as.numeric(Years), na.rm = T), 2)

# Distribution of age of studies
# MetaboLights
MLSYear <- as.data.frame(table(ML$Year))
colnames(MLSYear) <- c("Year", "Frequency")
MLSYear$Repository <- rep("MetaboLights")
# Metabolomics Workbench
MWSYear <- as.data.frame(table(MW$Year))
colnames(MWSYear) <- c("Year", "Frequency")
MWSYear$Repository <- rep("Metabolomics Workbench")
# GNPS
GNPSSYear <- as.data.frame(table(GNPS$Year))
colnames(GNPSSYear) <- c("Year", "Frequency")
GNPSSYear$Repository <- rep("GNPS")
# All studies
AllSYears <-  as.data.frame(table(Years))
colnames(AllSYears) <- c("Year", "Frequency")
AllSYears$Repository <- rep("All")

# Recombine data into single data frame
StudYear <- rbind(AllSYears, MLSYear, MWSYear, GNPSSYear)
StudYear$Repository <- as.factor(StudYear$Repository)

# Remove 2018, 2019 and 2050 for ploting
y2018 <- grepl("2018",StudYear$Year)
Till2017 <- StudYear[!y2018,]
y2019 <- grepl("2019",Till2017$Year)
Till2017 <- Till2017[!y2019,]
y2050 <- grepl("2050",Till2017$Year)
Till2017 <- Till2017[!y2050,]

# Add zero years for plotting
MW0 <- data.frame(Year = 2012, Frequency = 0, Repository = "Metabolomics Workbench")
GNPS0 <- data.frame(Year = 2013, Frequency = 0, Repository = "GNPS")

# recombine data frames
Till2017$Year <- factor(Till2017$Year, levels = c("2011",levels(Till2017$Year)))
Till2017 <- rbind(Till2017, MW0, GNPS0)
```

# Tables

## Table 3.7. Studies that reuse public available metabolomics data


```{r datalevels, results = 'asis'}
Paper <- DataReuse$Link
kable(cbind(DataReuse[,1:4], Paper), caption = "Studies that reuse public available metabolomics data, as of 15^th^ February 2018. The table shows articles that reuse publicly available data, the repository(-ies) that were the source of the data, the year the article was published and classification of how the data were re-used.") %>%
  kable_styling(full_width = F, bootstrap_options = c("hover", "responsive"))
```

# Figures
Code that was used to generate raw figures. Figures were further processed in Adobe Illustrator.

## Figure 3.10. The number of articles that reuse metabolomics data over time. 
Data reuse across all repositories is shown in red, GNPS reuse is shown in blue, MetaboLights reuse is shown in green and Metabolomics Workbench is shown in purple. Some articles reuse data from multiple repositories (both MetaboLights and Metabolomics Workbench), so the total number of articles that reuse metabolomics data per year is not the sum of the articles that reuse data from each repository per year. The launch year of each repository is also highlighted.

```{r reuseperyear, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(No2018, aes(Year, Frequency, color=Repository, group=Repository, shape=Repository))  + 
  geom_line() +
  geom_point(size = 3) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20)) +
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = 18), linetype="dashed", color = "black", size=0.2) +
  geom_segment(aes(x = 2, y = 0, xend = 2, yend = 17), linetype="dashed", color = "black", size=0.2) +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 18), linetype="dashed", color = "black", size=0.3) +
  theme_bw() +
  theme(
    legend.position="top",
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title.align=0.5) +
  scale_color_manual(values = c("#D55E00","#0072B2", "#009E73", "#CC79A7")) +
  guides(color = guide_legend(title.position = "top")) +
  annotate("text", x = 1, y = 19, label = "MetaboLights \n launched") +
  annotate("text", x = 2, y = 18.5, label = "Metabolomics \n Workbench \n launched") +
  annotate("text", x = 3, y = 19, label = "GNPS \n launched")
```

## Figure 3.11A. The percentage of studies reused at each frequency including all studies

The percentage of studies reused at each frequency: 0, 1, 2, 3, or 4 times, as of 15^th^ February 2018, including reuse in all studies.
```{r reuseperstudyall, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(Reuseperg, aes(x=Repository, y=Percentage, fill=Frequency))  +
  geom_bar(colour="black", stat="identity", width = 0.7) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 101)) +
  scale_fill_manual(values = c('#0868ac','#43a2ca','#7bccc4',"#bae4bc",'#f0f9e8')) +
  ylab("Frequency of Reuse per Study (%)")+
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        legend.text.align = 0,
        #legend.text = element_text(face = "italic"),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

## Figure 3.11B. The percentage of studies reused at each frequency excluding Spicer *et al.* (2017)

The percentage of studies reused at each frequency: 0, 1, 2, 3, or 4 times, as of 15^th^ February 2018 excluding reuse by Spicer \textit{et al.} (2017).

```{r reuseperstudynospicer, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(Reusepernospicerg, aes(x=Repository, y=Percentage, fill=Frequency))  +
  geom_bar(colour="black", stat="identity", width = 0.7) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 101)) +
  scale_fill_manual(values = c('#0868ac','#43a2ca','#7bccc4',"#bae4bc",'#f0f9e8')) +
  ylab("Frequency of Reuse per Study (%)")+
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        legend.text.align = 0,
        #legend.text = element_text(face = "italic"),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

## Figure 3.12A. The frequency of metabolomics studies released per year

The total number of studies released per year is shown in red, GNPS studies are in blue, MetaboLights are in green and Metabolomics Workbench are in purple.

```{r sAge, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(Till2017, aes(Year, Frequency, color=Repository, group=Repository, shape=Repository))  + 
  geom_line() +
  geom_point(size = 3) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1000)) +
  theme_bw() +
  theme(
    legend.position="top",
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title.align=0.5) +
  scale_color_manual(values = c("#D55E00","#0072B2", "#009E73", "#CC79A7")) +
  guides(color = guide_legend(title.position = "top"))
```


## Figure 3.12B. The frequency of publicly available studies and time until data reuse

```{r time2reuse, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(Time2Reuse, aes(Years, Percentage, group = 1))  + 
  geom_line() +
  geom_point() +
  ylab("Percentage of Reused Studies") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
  theme_bw() +
  theme(
    legend.position="top",
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title.align=0.5)
```


## Supplementary Figure 1. Types of Reuse

```{r reusetype, results = 'asis',  fig.align = "center"}
# Extract the frequency of types of reuse
TypeReuse <- as.data.frame(table(DataReuse$Classification))
colnames(TypeReuse) <- c("Classification", "Frequency")
TypeReuse$Percentage <- TypeReuse$Frequency/sum(TypeReuse$Frequency) * 100
TypeReuseOrdered <- TypeReuse[order(-TypeReuse$Percentage),]

colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)')

plot_ly(TypeReuseOrdered, labels = ~Classification, values = ~Percentage, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        text = ~paste(Frequency, "Studies"),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
        #The 'pull' attribute can also be used to create space between the sectors
        showlegend = FALSE) %>%
  layout(title = 'Study Classification',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```

## Supplementary Figure 2. The percentage of studies published per year that have been reused including all studies

```{r timeall, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(PerAge, aes(Year, Percentage, color=Repository, group=Repository, shape=Repository))  + 
  geom_line() +
  geom_point(size = 3) +
  ylab("Reused Studies (%)") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = 49), linetype="dashed", color = "black", size=0.3) +
  geom_segment(aes(x = 2, y = 0, xend = 2, yend = 49), linetype="dashed", color = "black", size=0.15) +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 22), linetype="dashed", color = "black", size=0.35) +
  theme_bw() +
  theme(
    legend.position="top",
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title.align=0.5) +
  scale_color_manual(values = c("#0072B2", "#009E73", "#CC79A7")) +
  guides(color = guide_legend(title.position = "top")) +
  annotate("text", x = 1.5, y = 45, label = "MetaboLights \n launched") +
  annotate("text", x = 2.55, y = 42.5, label = "Metabolomics \n Workbench \n launched") +
  annotate("text", x = 3.35, y = 18, label = "GNPS \n launched")
```

## Supplementary Figure 3. The percentage of studies published per year that have been reused excluding Spicer *et al.* (2017)

```{r timenospicer, results = 'asis', fig.height = 5, fig.width = 7, fig.align = "center"}
ggplot(PerNSAge, aes(Year, Percentage, color=Repository, group=Repository, shape=Repository))  + 
  geom_line() +
  geom_point(size = 3) +
  ylab("Reused Studies (%)") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = 49), linetype="dashed", color = "black", size=0.3) +
  geom_segment(aes(x = 2, y = 0, xend = 2, yend = 6), linetype="dashed", color = "black", size=0.1) +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 22), linetype="dashed", color = "black", size=0.35) +
  theme_bw() +
  theme(
    legend.position="top",
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title.align=0.5) +
  scale_color_manual(values = c("#0072B2", "#009E73", "#CC79A7")) +
  guides(color = guide_legend(title.position = "top")) +
  annotate("text", x = 1, y = 56, label = "MetaboLights \n launched") +
  annotate("text", x = 2, y = 16, label = "Metabolomics \n Workbench \n launched") +
  annotate("text", x = 3, y = 30, label = "GNPS \n launched")
```

## Supplementary Table 1. The frequency of reuse per study
```{r reusepers, results = 'asis'}
ReuseStud <- UniqueStudiesCount(DataReuse)
colnames(ReuseStud) <- c("Study", "Frequency")
kable(ReuseStud, row.names = F) %>%
  kable_styling(full_width = F)
```